---
title: "BIS15L_AI_class"
author: "Amina Muhic"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?
AI, or artificial intelligence, refers to the simulation of human intelligence in machines that are programmed to think and mimic human actions. These machines are designed to perform tasks that typically require human intelligence, such as visual perception, speech recognition, decision-making, and language translation. AI systems use techniques like machine learning, neural networks, natural language processing, and computer vision to learn from data, recognize patterns, and make decisions. AI is used in various fields, including healthcare, finance, transportation, entertainment, and more, to automate processes, improve efficiency, and enhance decision-making.

2. What is ChatGPT?
ChatGPT is the name of the conversational AI system you're currently interacting with. It's part of the GPT (Generative Pre-trained Transformer) family of language models developed by OpenAI. ChatGPT is designed to understand and generate human-like text based on the input it receives. It's trained on a vast amount of text from the internet and other sources to understand and generate responses in a wide range of topics and conversational contexts.

3. What is Github Copilot?
GitHub Copilot is an AI-powered code completion tool developed by GitHub in collaboration with OpenAI. It assists developers by suggesting code snippets and completing lines of code as they type. Copilot is trained on a large dataset of code from GitHub repositories and uses machine learning models to understand context and generate relevant code suggestions. It supports multiple programming languages and aims to speed up the coding process, improve code quality, and help developers discover new techniques and patterns.

4. [Try it yourself: Insert your own question here]
How is ChatGPT different from GitHub Copilot?

ChatGPT and GitHub Copilot are both AI-powered tools, but they serve different purposes and operate in different domains:

ChatGPT: ChatGPT is a conversational AI model designed for generating human-like text in response to user input. It's trained on a diverse range of text from the internet and can engage in conversations, answer questions, provide explanations, and more. ChatGPT is used primarily for natural language understanding and generation tasks, such as chatbots, language translation, and content generation.

GitHub Copilot: GitHub Copilot is an AI-powered code completion tool developed by GitHub in collaboration with OpenAI. It assists developers by suggesting code snippets and completing lines of code as they type. Copilot is trained on a large dataset of code from GitHub repositories and uses machine learning models to understand context and generate relevant code suggestions. It's used to speed up the coding process, improve code quality, and help developers discover new techniques and patterns.

In summary, ChatGPT focuses on natural language understanding and generation, while GitHub Copilot focuses on code completion and generation specifically for software development tasks.


## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix

netflix <- readr::read_csv("data/netflix_titles.csv")
```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
head(netflix)

```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
# use filter to sort out the movies from the netflix data
```


Which director has the most movies on netflix?
```{r}
# use count to get the director with the most movies on netflix
```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*
```{r}
# use filter to sort out the directors with the most movies but remove the NAs
```

[insert your prompt here]
in ChatGPT: under a variable in a data frame, how would you separate an observation that contains two observations separated by a comma in R?
```{r}
netflix %>% 
  separate_rows(director, sep = ',')
```



Which director had the most movies by release year and remove the NA's?
```{r}

```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


[Insert your prompt here]
```{r}

```



Make a bar plot of the top directors with the most movies on netflix
```{r}


```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*



[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt


*Now use ChatGPT, Copilot or previous notes to fix the code*





## That's it! Let's take a break and then move on to working on group projects.
