---
output: 
  html_document: 
    keep_md: yes
---
## Lab 6 Warmup

library calls
```{r}
library(tidyverse)
library(janitor)
```

1. load the `bison.csv` data
```{r}
bison <- readr::read_csv("data/bison.csv")
```

2. what are the dimesions and structure of the data?
```{r}
dim(bison)
```

```{r}
glimpse(bison)
```

3. we are only interested in code, sex, weight, year of birth. restrict the data to these variables and store the dataframe as a new object.
```{r}
bison_new <- bison %>% 
  select("data_code", "animal_sex", "animal_weight", "animal_yob")
```

4. pull out the animals born between 1980 and 1990
```{r}
bison_new %>% 
  filter(animal_yob %in% c(1980, 1990))
```

5. how many male and female bison are represented between 1980 and 1990?
```{r}
bison_aged <- bison_new %>% 
  filter(animal_yob %in% c(1980, 1990))
table(bison_aged$animal_sex)
```

6. between 1980-1990, were males or females larger on average?
```{r}
bison_aged_f <- filter(bison_aged, animal_sex == "F")
mean(bison_aged_f$animal_weight, na.rm = T)
```

```{r}
bison_aged_m <- filter(bison_aged, animal_sex == "M")
mean(bison_aged_m$animal_weight, na.rm = T)
```
Males are larger on average between 1980-1990.
